---
title: "Premiere League Analysis"
author: "Kaan Bilgin"
date: "11/20/2020"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup,include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# 1.Introduction

Sports forecasting is important for sports fans, team managers, sponsors, the media and the growing
number of punters who bet on online platforms. Widespread demand for professional advice regarding
the results of sporting events is met by a variety of expert forecasts, usually in the form of
recommendations from tipsters. In addition, betting odds offer a type of predictor and source of expert
advice regarding sports outcomes. Whereas fixed odds reflect the (expert) predictions of bookmakers,
the odds in pari-mutuel betting markets indicate the combined expectations of all punters, which
implies an aggregated expert prediction. Expert forecasts of sport outcomes often come from so-called ‘tipsters’, whose predictions appear in
sports journals or daily newspapers. Tipsters are usually independent experts who do not apply a formal
model but rather derive their predictions from their experience or intuition. They generally provide
forecasts for only a specific selection of games, often related to betting. No immediate financial
consequences result from the predictions of tipsters. Empirical evidence regarding the forecast
accuracy of tipsters shows that their ability is limited. This homework is about understanding the behaviour of different betting companies and leagues with
the use of available information from different sources (odds from different betting companies, team
status and etc.).

As background for the rest of the formulas and assumptions, please check [this](http://www.it.uu.se/edu/course/homepage/projektTDB/ht15/project16/Project16_Report.pdf) paper by Mirza and Fejes, also statistics of the Premier League seasons provided by [this](http://betamatics.com/) site.

# 2.Libraries

Importing libraries. They mainly used for data handling and plotting.

```{r library, message=FALSE, warning=FALSE}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(gridExtra)
library(data.table)
```

For this analysis English Premier League 2018-2019, 2019-2020 and 2020-2021 season data are used. Data imported like following;

data18_19 contains 2018-2019 season's data
data19_20 contains 2019-2020 season's data
data20_21 contains 2020-2021 season's data
data_league contains all of the data.

```{r import data, message=FALSE, warning=FALSE}

data18_19 = fread("C:/users/Kaan/Desktop/Master/Data_Mining/HW1/18-19.CSV", select = 2:56)
data19_20 = fread("C:/users/Kaan/Desktop/Master/Data_Mining/HW1/19-20.csv", select = 2:56)
data20_21 = fread("C:/users/Kaan/Desktop/Master/Data_Mining/HW1/20-21.csv", select = 2:56)

data_league <- bind_rows(data18_19, data19_20, data20_21)


```

# 3.Analysis and Plots



## 3.1.Task 1

### 3.1.1.Task A: Histogram for Home Goals

```{r plots, message=FALSE, warning=FALSE}
ggplot(data=data_league, aes(x=FTHG), binwidth=1) + geom_histogram(col="black", fill="Cyan", alpha = .5) +labs(title="Histogram for Home Goals", x="Home Goals", y="Number of Games")
  

```
### 3.1.2.Task B: Histogram for Away Goals

```{r plots_2, message=FALSE, warning=FALSE}
ggplot(data=data_league, aes(x=FTAG), binwidth=1) + geom_histogram(col="black",fill="dark blue", alpha = .5) + 
  labs(title="Histogram for Away Goals", x="Away Goals", y="Number of Games") 
```
### 3.1.3.Task C: Histogram for Home - Away Goals

```{r plots_3, message=FALSE, warning=FALSE}
ggplot(data_league) + geom_histogram(aes(x = FTHG-FTAG), binwidth=1, col="black", fill="orange") + xlab("Home Goals – Away Goals") + ylab("Number of Games") + theme(plot.caption=element_text(hjust = 0.5))
```

#### 3.1.3.1.Task C Bonus: Poisson Distribution Similarity Check


```{r plots_4, message=FALSE, warning=FALSE}
#sample mean calculations and lenght

lambda_hg= mean(data_league$FTHG,trim=0,na.rm=T)
lambda_ag= mean(data_league$FTAG,trim=0,na.rm=T)
n = length(data_league$FTHG)


# poisson density - home

p1=ggplot(data_league,aes(FTHG)) + geom_histogram(aes(fill="Home Goals"),binwidth = 1,col="black",show.legend = FALSE)+
  geom_line(aes(y=dpois(x=FTHG,lambda = lambda_hg)*n)) +
  labs(x="Home Goals",y="Number of Games",title=paste0("Poisson with Lambda: ",round(lambda_hg,2)))

# poisson density - away

p2=ggplot(data_league,aes(FTAG)) + geom_histogram(aes(fill="Away Goals"),binwidth = 1,col="black",show.legend = FALSE)+
  geom_line(aes(y=dpois(x=FTAG,lambda = lambda_hg)*n)) +
  labs(x="Away Goals",y="Number of Games",title=paste0("Poisson with Lambda: ",round(lambda_ag,2)))

# combine home and away plots
grid.arrange(p1,p2,ncol=2)
```

## 3.2.Task 2: Analysis of Bookmakers Draw Bet Odds 

For this task I choose (B365H, B365D, B365A) , (BWH, BWD, BWA) , (IWH, IWD, IWA) and (PSH, PSD, PSA) as Bookmakers. 

Let's calculate the P(Home Win), P(Draw) and P(Away Win) for selected bet options:
```{r setup_1, message=FALSE, warning=FALSE}
data_league$prob_B365H <- 1/(data_league$B365H)
data_league$prob_B365D <- 1/(data_league$B365D)
data_league$prob_B365A <- 1/(data_league$B365A)

data_league$prob_BWH <- 1/(data_league$BWH)
data_league$prob_BWD <- 1/(data_league$BWD)
data_league$prob_BWA <- 1/(data_league$BWA)

data_league$prob_IWH <- 1/(data_league$IWH)
data_league$prob_IWD <- 1/(data_league$IWD)
data_league$prob_IWA <- 1/(data_league$IWA)

data_league$prob_PSH <- 1/(data_league$PSH)
data_league$prob_PSA <- 1/(data_league$PSA)
data_league$prob_PSD <- 1/(data_league$PSD)
```

```{r setup_2, message=FALSE, warning=FALSE}

data_league %>%
  colnames()
```

Normalization of the previous odds:

```{r setup_3, message=FALSE, warning=FALSE}

#Normalized B365:
data_league$N_prob_B365H <- data_league$prob_B365H/(data_league$prob_B365H + data_league$prob_B365D + data_league$prob_B365A)
data_league$N_prob_B365D <- data_league$prob_B365D/(data_league$prob_B365H + data_league$prob_B365D + data_league$prob_B365A)
data_league$N_prob_B365A <- data_league$prob_B365A/(data_league$prob_B365H + data_league$prob_B365D + data_league$prob_B365A)

#Normalized BW:
data_league$N_prob_BWH <- data_league$prob_BWH/(data_league$prob_BWH + data_league$prob_BWD + data_league$prob_BWA)
data_league$N_prob_BWD <- data_league$prob_BWD/(data_league$prob_BWH + data_league$prob_BWD + data_league$prob_BWA)
data_league$N_prob_BWA <- data_league$prob_BWA/(data_league$prob_BWH + data_league$prob_BWD + data_league$prob_BWA)

#Normalized IW:
data_league$N_prob_IWH <- data_league$prob_IWH/(data_league$prob_IWH + data_league$prob_IWD + data_league$prob_IWA)
data_league$N_prob_IWD <- data_league$prob_IWD/(data_league$prob_IWH + data_league$prob_IWD + data_league$prob_IWA)
data_league$N_prob_IWA <- data_league$prob_IWA/(data_league$prob_IWH + data_league$prob_IWD + data_league$prob_IWA)

#Normalized PS:
data_league$N_prob_PSH <- data_league$prob_PSH/(data_league$prob_PSH + data_league$prob_PSA + data_league$prob_PSD)
data_league$N_prob_PSA <- data_league$prob_PSA/(data_league$prob_PSH + data_league$prob_PSA + data_league$prob_PSD)
data_league$N_prob_PSD <- data_league$prob_PSD/(data_league$prob_PSH + data_league$prob_PSA + data_league$prob_PSD)
```

```{r setup_4, message=FALSE, warning=FALSE}
data_league %>%
  colnames()
```

Discretization and Plotting for Task 2.3:

BET 365:
```{r setup_5, message=FALSE, warning=FALSE}
#FTR
data_league = data_league %>%
  mutate(FTR_tie = ifelse(FTR == 'D', 1, 0)) %>%
  mutate(bins = case_when( N_prob_B365H - N_prob_B365A < -0.9  ~ "1",
                           N_prob_B365H - N_prob_B365A < -0.8  ~ "2",
                           N_prob_B365H - N_prob_B365A < -0.7  ~ "3",
                           N_prob_B365H - N_prob_B365A < -0.6  ~ "4",
                           N_prob_B365H - N_prob_B365A < -0.5  ~ "5",
                           N_prob_B365H - N_prob_B365A < -0.4  ~ "6",
                           N_prob_B365H - N_prob_B365A < -0.3  ~ "7",
                           N_prob_B365H - N_prob_B365A < -0.2  ~ "8",
                           N_prob_B365H - N_prob_B365A < -0.1  ~ "9",
                           N_prob_B365H - N_prob_B365A <  0    ~ "10",
                           N_prob_B365H - N_prob_B365A <  0.1  ~ "11",
                           N_prob_B365H - N_prob_B365A <  0.2  ~ "12",
                           N_prob_B365H - N_prob_B365A <  0.3  ~ "13",
                           N_prob_B365H - N_prob_B365A <  0.4  ~ "14",
                           N_prob_B365H - N_prob_B365A <  0.5  ~ "15",
                           N_prob_B365H - N_prob_B365A <  0.6  ~ "16",
                           N_prob_B365H - N_prob_B365A <  0.7  ~ "17",
                           N_prob_B365H - N_prob_B365A <  0.8  ~ "18",
                           N_prob_B365H - N_prob_B365A <  0.9  ~ "19",
                           TRUE ~ "20")) %>%
  group_by(bins) %>%
  mutate(drawB365 = mean(FTR_tie))

p3 <- ggplot() + geom_point(aes(x = N_prob_B365H - N_prob_B365A, y = drawB365), data = data_league, color = "red") +geom_point(aes(x = N_prob_B365H - N_prob_B365A, y = N_prob_B365D), data = data_league, color = "darkgray") +
  labs(title = 'Bet365',
       x = 'P(Home) - P(Away) ',
       y = 'P(Draw) ')
p3

```

For BW:

```{r setup_6, message=FALSE, warning=FALSE}
#BW
data_league = data_league %>%
  mutate(FTR_tie = ifelse(FTR == 'D', 1, 0)) %>%
  mutate(bins = case_when( N_prob_BWH - N_prob_BWA < -0.9  ~ "1",
                           N_prob_BWH - N_prob_BWA < -0.8  ~ "2",
                           N_prob_BWH - N_prob_BWA < -0.7  ~ "3",
                           N_prob_BWH - N_prob_BWA < -0.6  ~ "4",
                           N_prob_BWH - N_prob_BWA < -0.5  ~ "5",
                           N_prob_BWH - N_prob_BWA < -0.4  ~ "6",
                           N_prob_BWH - N_prob_BWA < -0.3  ~ "7",
                           N_prob_BWH - N_prob_BWA < -0.2  ~ "8",
                           N_prob_BWH - N_prob_BWA < -0.1  ~ "9",
                           N_prob_BWH - N_prob_BWA <  0    ~ "10",
                           N_prob_BWH - N_prob_BWA <  0.1  ~ "11",
                           N_prob_BWH - N_prob_BWA <  0.2  ~ "12",
                           N_prob_BWH - N_prob_BWA <  0.3  ~ "13",
                           N_prob_BWH - N_prob_BWA <  0.4  ~ "14",
                           N_prob_BWH - N_prob_BWA <  0.5  ~ "15",
                           N_prob_BWH - N_prob_BWA <  0.6  ~ "16",
                           N_prob_BWH - N_prob_BWA <  0.7  ~ "17",
                           N_prob_BWH - N_prob_BWA <  0.8  ~ "18",
                           N_prob_BWH - N_prob_BWA <  0.9  ~ "19",
                           TRUE ~ "20")) %>%
  group_by(bins) %>%
  mutate(drawBW = mean(FTR_tie))

p4 <- ggplot() + geom_point(aes(x = N_prob_BWH - N_prob_BWA, y = drawBW), data = data_league, color = "orange2") +geom_point(aes(x = N_prob_BWH - N_prob_BWA, y = N_prob_BWD), data = data_league, color = "darkgray") +
  labs(title = 'BW',
       x = 'P(Home) - P(Away) ',
       y = 'P(Draw) ')

p4
```

For IW: 

```{r setup_7, message=FALSE, warning=FALSE}
#For IW:
data_league = data_league %>%
  mutate(FTR_tie = ifelse(FTR == 'D', 1, 0)) %>%
  mutate(bins = case_when( N_prob_IWH - N_prob_IWA < -0.9  ~ "1",
                           N_prob_IWH - N_prob_IWA < -0.8  ~ "2",
                           N_prob_IWH - N_prob_IWA < -0.7  ~ "3",
                           N_prob_IWH - N_prob_IWA < -0.6  ~ "4",
                           N_prob_IWH - N_prob_IWA < -0.5  ~ "5",
                           N_prob_IWH - N_prob_IWA < -0.4  ~ "6",
                           N_prob_IWH - N_prob_IWA < -0.3  ~ "7",
                           N_prob_IWH - N_prob_IWA < -0.2  ~ "8",
                           N_prob_IWH - N_prob_IWA < -0.1  ~ "9",
                           N_prob_IWH - N_prob_IWA <  0    ~ "10",
                           N_prob_IWH - N_prob_IWA <  0.1  ~ "11",
                           N_prob_IWH - N_prob_IWA <  0.2  ~ "12",
                           N_prob_IWH - N_prob_IWA <  0.3  ~ "13",
                           N_prob_IWH - N_prob_IWA <  0.4  ~ "14",
                           N_prob_IWH - N_prob_IWA <  0.5  ~ "15",
                           N_prob_IWH - N_prob_IWA <  0.6  ~ "16",
                           N_prob_IWH - N_prob_IWA <  0.7  ~ "17",
                           N_prob_IWH - N_prob_IWA <  0.8  ~ "18",
                           N_prob_IWH - N_prob_IWA <  0.9  ~ "19",
                           TRUE ~ "20")) %>%
  group_by(bins) %>%
  mutate(drawIW = mean(FTR_tie))

p5 <- ggplot() + geom_point(aes(x = N_prob_IWH - N_prob_IWA, y = drawIW), data = data_league, color = "orange") +geom_point(aes(x = N_prob_IWH - N_prob_IWA, y = N_prob_IWD), data = data_league, color = "darkgray") +
  labs(title = 'IW',
       x = 'P(Home) - P(Away) ',
       y = 'P(Draw) ')
p5

```

For PS:
```{r setup_8, message=FALSE, warning=FALSE}
#For PS:
data_league = data_league %>%
  mutate(FTR_tie = ifelse(FTR == 'D', 1, 0)) %>%
  mutate(bins = case_when( N_prob_PSH - N_prob_PSA < -0.9  ~ "1",
                           N_prob_PSH - N_prob_PSA < -0.8  ~ "2",
                           N_prob_PSH - N_prob_PSA < -0.7  ~ "3",
                           N_prob_PSH - N_prob_PSA < -0.6  ~ "4",
                           N_prob_PSH - N_prob_PSA < -0.5  ~ "5",
                           N_prob_PSH - N_prob_PSA < -0.4  ~ "6",
                           N_prob_PSH - N_prob_PSA < -0.3  ~ "7",
                           N_prob_PSH - N_prob_PSA < -0.2  ~ "8",
                           N_prob_PSH - N_prob_PSA < -0.1  ~ "9",
                           N_prob_PSH - N_prob_PSA <  0    ~ "10",
                           N_prob_PSH - N_prob_PSA <  0.1  ~ "11",
                           N_prob_PSH - N_prob_PSA <  0.2  ~ "12",
                           N_prob_PSH - N_prob_PSA <  0.3  ~ "13",
                           N_prob_PSH - N_prob_PSA <  0.4  ~ "14",
                           N_prob_PSH - N_prob_PSA <  0.5  ~ "15",
                           N_prob_PSH - N_prob_PSA <  0.6  ~ "16",
                           N_prob_PSH - N_prob_PSA <  0.7  ~ "17",
                           N_prob_PSH - N_prob_PSA <  0.8  ~ "18",
                           N_prob_PSH - N_prob_PSA <  0.9  ~ "19",
                           TRUE ~ "20")) %>%
  group_by(bins) %>%
  mutate(drawPS = mean(FTR_tie))

p6 <- ggplot() + geom_point(aes(x = N_prob_PSH - N_prob_PSA, y = drawPS), data = data_league, color = "red2") +geom_point(aes(x = N_prob_PSH - N_prob_PSA, y = N_prob_PSD), data = data_league, color = "darkgray") +
  labs(title = 'PS',
       x = 'P(Home) - P(Away) ',
       y = 'P(Draw) ')
p6

grid.arrange(p3,p4,p5,p6,ncol=2)

```

### 3.2.1.Results and Discussion
4 Bookmakers are selected and the results shows us there is a bias between calculated draw probability and actual draw probability. 


## 3.3.Task 3: Effect of the Noise 

Effect of the noise analyzed in this section. Noise in this analyze selected as Red Card.

```{r setup_9, message=FALSE, warning=FALSE}
#Red Card for B365:
data_RED = data_league %>%
  filter(HR + AR == 0)

data_RED = data_RED %>%
  mutate(FTR_tie = ifelse(FTR == 'D', 1, 0)) %>%
  mutate(bins = case_when( N_prob_B365H - N_prob_B365A < -0.9  ~ "1",
                           N_prob_B365H - N_prob_B365A < -0.8  ~ "2",
                           N_prob_B365H - N_prob_B365A < -0.7  ~ "3",
                           N_prob_B365H - N_prob_B365A < -0.6  ~ "4",
                           N_prob_B365H - N_prob_B365A < -0.5  ~ "5",
                           N_prob_B365H - N_prob_B365A < -0.4  ~ "6",
                           N_prob_B365H - N_prob_B365A < -0.3  ~ "7",
                           N_prob_B365H - N_prob_B365A < -0.2  ~ "8",
                           N_prob_B365H - N_prob_B365A < -0.1  ~ "9",
                           N_prob_B365H - N_prob_B365A <  0    ~ "10",
                           N_prob_B365H - N_prob_B365A <  0.1  ~ "11",
                           N_prob_B365H - N_prob_B365A <  0.2  ~ "12",
                           N_prob_B365H - N_prob_B365A <  0.3  ~ "13",
                           N_prob_B365H - N_prob_B365A <  0.4  ~ "14",
                           N_prob_B365H - N_prob_B365A <  0.5  ~ "15",
                           N_prob_B365H - N_prob_B365A <  0.6  ~ "16",
                           N_prob_B365H - N_prob_B365A <  0.7  ~ "17",
                           N_prob_B365H - N_prob_B365A <  0.8  ~ "18",
                           N_prob_B365H - N_prob_B365A <  0.9  ~ "19",
                           TRUE ~ "20")) %>%
  group_by(bins) %>%
  mutate(drawB365 = mean(FTR_tie))

p7 <- ggplot() + geom_point(aes(x = N_prob_B365H - N_prob_B365A, y = drawB365), data = data_RED, color = "red") +geom_point(aes(x = N_prob_B365H - N_prob_B365A, y = N_prob_B365D), data = data_RED, color = "blue2") +
  labs(title = 'Bet365',
       x = 'P(Home) - P(Away) ',
       y = 'P(Draw) ')
  p7
# Comparison between Normal version vs No Red Card version  
grid.arrange(p3,p7,ncol=2)
```
For BW:
```{r setup_10, message=FALSE, warning=FALSE}
data_RED = data_league %>%
  filter(HR + AR == 0)

data_RED = data_RED %>%
  mutate(FTR_tie = ifelse(FTR == 'D', 1, 0)) %>%
  mutate(bins = case_when( N_prob_BWH - N_prob_BWA < -0.9  ~ "1",
                           N_prob_BWH - N_prob_BWA < -0.8  ~ "2",
                           N_prob_BWH - N_prob_BWA < -0.7  ~ "3",
                           N_prob_BWH - N_prob_BWA < -0.6  ~ "4",
                           N_prob_BWH - N_prob_BWA < -0.5  ~ "5",
                           N_prob_BWH - N_prob_BWA < -0.4  ~ "6",
                           N_prob_BWH - N_prob_BWA < -0.3  ~ "7",
                           N_prob_BWH - N_prob_BWA < -0.2  ~ "8",
                           N_prob_BWH - N_prob_BWA < -0.1  ~ "9",
                           N_prob_BWH - N_prob_BWA <  0    ~ "10",
                           N_prob_BWH - N_prob_BWA <  0.1  ~ "11",
                           N_prob_BWH - N_prob_BWA <  0.2  ~ "12",
                           N_prob_BWH - N_prob_BWA <  0.3  ~ "13",
                           N_prob_BWH - N_prob_BWA <  0.4  ~ "14",
                           N_prob_BWH - N_prob_BWA <  0.5  ~ "15",
                           N_prob_BWH - N_prob_BWA <  0.6  ~ "16",
                           N_prob_BWH - N_prob_BWA <  0.7  ~ "17",
                           N_prob_BWH - N_prob_BWA <  0.8  ~ "18",
                           N_prob_BWH - N_prob_BWA <  0.9  ~ "19",
                           TRUE ~ "20")) %>%
  group_by(bins) %>%
  mutate(drawBW = mean(FTR_tie))

p8 <- ggplot() + geom_point(aes(x = N_prob_BWH - N_prob_BWA, y = drawBW), data = data_RED, color = "orange2") +geom_point(aes(x = N_prob_BWH - N_prob_BWA, y = N_prob_BWD), data = data_RED, color = "blue2") +
  labs(title = 'BW',
       x = 'P(Home) - P(Away) ',
       y = 'P(Draw) ')
# Comparison between Normal Version vs No Red Card Version
grid.arrange(p4,p8,ncol=2)
```
For IW:
```{r setup_11, message=FALSE, warning=FALSE}
data_RED = data_league %>%
  filter(HR + AR == 0)

data_RED = data_RED %>%
  mutate(FTR_tie = ifelse(FTR == 'D', 1, 0)) %>%
  mutate(bins = case_when( N_prob_IWH - N_prob_IWA < -0.9  ~ "1",
                           N_prob_IWH - N_prob_IWA < -0.8  ~ "2",
                           N_prob_IWH - N_prob_IWA < -0.7  ~ "3",
                           N_prob_IWH - N_prob_IWA < -0.6  ~ "4",
                           N_prob_IWH - N_prob_IWA < -0.5  ~ "5",
                           N_prob_IWH - N_prob_IWA < -0.4  ~ "6",
                           N_prob_IWH - N_prob_IWA < -0.3  ~ "7",
                           N_prob_IWH - N_prob_IWA < -0.2  ~ "8",
                           N_prob_IWH - N_prob_IWA < -0.1  ~ "9",
                           N_prob_IWH - N_prob_IWA <  0    ~ "10",
                           N_prob_IWH - N_prob_IWA <  0.1  ~ "11",
                           N_prob_IWH - N_prob_IWA <  0.2  ~ "12",
                           N_prob_IWH - N_prob_IWA <  0.3  ~ "13",
                           N_prob_IWH - N_prob_IWA <  0.4  ~ "14",
                           N_prob_IWH - N_prob_IWA <  0.5  ~ "15",
                           N_prob_IWH - N_prob_IWA <  0.6  ~ "16",
                           N_prob_IWH - N_prob_IWA <  0.7  ~ "17",
                           N_prob_IWH - N_prob_IWA <  0.8  ~ "18",
                           N_prob_IWH - N_prob_IWA <  0.9  ~ "19",
                           TRUE ~ "20")) %>%
  group_by(bins) %>%
  mutate(drawIW = mean(FTR_tie))

p9 <- ggplot() + geom_point(aes(x = N_prob_IWH - N_prob_IWA, y = drawIW), data = data_RED, color = "orange") +geom_point(aes(x = N_prob_IWH - N_prob_IWA, y = N_prob_IWD), data = data_RED, color = "blue2") +
  labs(title = 'IW',
       x = 'P(Home) - P(Away) ',
       y = 'P(Draw) ')
# Comparison between Normal Version vs No Red Card Version
grid.arrange(p5,p9,ncol=2)
```
For PS:
```{r setup_12, message=FALSE, warning=FALSE}
data_RED = data_league %>%
  filter(HR + AR == 0)

data_RED = data_RED %>%
  mutate(FTR_tie = ifelse(FTR == 'D', 1, 0)) %>%
  mutate(bins = case_when( N_prob_PSH - N_prob_PSA < -0.9  ~ "1",
                           N_prob_PSH - N_prob_PSA < -0.8  ~ "2",
                           N_prob_PSH - N_prob_PSA < -0.7  ~ "3",
                           N_prob_PSH - N_prob_PSA < -0.6  ~ "4",
                           N_prob_PSH - N_prob_PSA < -0.5  ~ "5",
                           N_prob_PSH - N_prob_PSA < -0.4  ~ "6",
                           N_prob_PSH - N_prob_PSA < -0.3  ~ "7",
                           N_prob_PSH - N_prob_PSA < -0.2  ~ "8",
                           N_prob_PSH - N_prob_PSA < -0.1  ~ "9",
                           N_prob_PSH - N_prob_PSA <  0    ~ "10",
                           N_prob_PSH - N_prob_PSA <  0.1  ~ "11",
                           N_prob_PSH - N_prob_PSA <  0.2  ~ "12",
                           N_prob_PSH - N_prob_PSA <  0.3  ~ "13",
                           N_prob_PSH - N_prob_PSA <  0.4  ~ "14",
                           N_prob_PSH - N_prob_PSA <  0.5  ~ "15",
                           N_prob_PSH - N_prob_PSA <  0.6  ~ "16",
                           N_prob_PSH - N_prob_PSA <  0.7  ~ "17",
                           N_prob_PSH - N_prob_PSA <  0.8  ~ "18",
                           N_prob_PSH - N_prob_PSA <  0.9  ~ "19",
                           TRUE ~ "20")) %>%
  group_by(bins) %>%
  mutate(drawPS = mean(FTR_tie))

p10 <- ggplot() + geom_point(aes(x = N_prob_PSH - N_prob_PSA, y = drawPS), data = data_RED, color = "red2") +geom_point(aes(x = N_prob_PSH - N_prob_PSA, y = N_prob_PSD), data = data_RED, color = "blue2") +
  labs(title = 'PS',
       x = 'P(Home) - P(Away) ',
       y = 'P(Draw) ')
# Comparison between Normal Version vs No Red Card Version
grid.arrange(p6,p10,ncol=2)
```

```{r setup_18, message=FALSE, warning=FALSE}

# total plots for no red cards
grid.arrange(p7,p8,p9,p10,ncol=2)
```

### 3.3.1.Results and Discussion

Generally minimum and maximum calculated draw probabilities are changed. 


Press [here](https://bu-ie-582.github.io/fall20-kaanblgn/) to return to homepage.